The Dynamics of Anthropomorphic Effects in Human-Robot Interaction

### 2014-03-26 paper.tex
anthropomorphic design section DONE (almost)

Along this line, this article aims at fostering a better understanding of
anthropomorphism (noted \Ant in this paper) by considering the phenomenon
\textbf{as a whole}: robotics researchers often tend to believe that
anthropomorphism only describes a static set of human-like features of a robot
(like shape, speech capabilities, facial expressions, etc.). We refer to these
characteristics as the \emph{anthropomorphic design} of the
robot~\cite{fink_anthropomorphism_2012}. But \emph{anthropomorphism} in general
refers to the larger \emph{social phenomenon} that emerges from the (real or
imagined) \emph{interaction} between a robot and a
user~\citep{persson_anthropomorphism_2000}. According to~\cite{epley_when_2008},
this includes for instance emotional states, motivations, intentions
\emph{ascribed by the user} to the robot. As such, the dynamic and
socio-cognitive dimensions of anthropomorphism are essential to understand the
complex bonds that human users build with the robots.

@@

On one hand, developers and designers have discovered that human-like robots may better 
match the affordances that this to humans adapted world brings along. 
In general, it is suggested that increasing the human-likeness of a robot (up to a 
certain threshold) will lead to 
more perceived \textit{affinity} from the human side \citep{mori_uncanny_1970} (see 
Figure \ref{fig:uncanny_valley}).
This further increases anthropomorphism, and as such
appears to be beneficial for human-robot relations, especially with socially 
interactive robots \cite{fong_survey_2003}. Consequently, the domain of (social) 
robotics tries to exploit this by using
anthropomorphic forms in robots.
Humans seem to readily accept and prefer the 
anthropomorphic form of a robot (up to a certain point), when compared to purely mechanical 
robots. A possible explanation for this is that perceiving a system as
human-like (or also pet-like) compared to machine-like changes the 
way how users try to understand and make sense of this system. The reason for this is 
because humans tend to project everyday 
\hl{expectations} of human and animalistic life onto artificial agents that use 
anthropomorphic (or zoomorphic) forms \cite{schmitz_concepts_2011}. 
Both cognitive science 
and social psychology can provide explanations for this process of projection. The basic idea 
is that \textit{"the more similar in 
appearance, the more people are likely to use themselves as a source of induction 
and anthropomorphize these nonhuman agents"} \citep{epley_seeing_2007}. Studies have further 
shown that perceived similarity between the user and the robot influences the intensity of 
the participants' attributions towards the robot. Thus,the anthropomorphic design of a robot 
impacts the user's perception of the robot, and how far the user is likely to attribute and 
experience qualities such as intelligence, fun, or intentions in the robot.
(We will come back to this more thoughtful reaction to anthropomorphic forms, in more detail 
later, in Section \ref{sec:human-centered}).

In short, the central point for the design of sociable robots is that 
humans tend to apply the same communication references to anthropomorphic robots 
that they would use with other humans, which in turn enhances feeling familiar and 
comfortable when interacting with the otherwise unknown system.
This positive effect has been shown in various 
empirical studies. It has been found that robots with human-like forms can enhance social 
responses from humans which in turn can positively impact on acceptance 
\citep{duffy_anthropomorphism_2003,goetz_cooperation_2002,venkatesh_theoretical_2000}. 
People responded more positively to an artifact that displayed human-like 
behavioral characteristics (emotions, facial expression) in contrast to a purely 
functional design 
\citep{eyssel_anthropomorphic_2010,krach_can_2008,reeves_media_1996,riek_how_2009}. 
Overall, these results suggest that human-like features in robots enhance 
anthropomorphism, and in turn create a social view toward the robot.

@@

Mental models have also been studied in HRI. Starting point for this research is the assumption that if a system behaves much like a human being (\eg a robot that emits a human voice), people's mental model of the system's behavior may approach their mental model of humans (though the model may differ in some important aspects). In turn, people's estimation of an agent's knowledge model and its capabilities affects the way they relate to it. This holds important implications for the resulting interaction with the system, the user experience, and the acceptance. Previous research examined the validity of the mental model concept with various kinds of robots \cite{schmitz_concepts_2011,kiesler_mental_2002}. Findings suggest that people tend to hold richer mental models of anthropomorphic robots in contrast to mechanic ones \cite{kiesler_mental_2002}.\footnote{\hl{Later in this article, we may however come to question how far this result is based on expectations from the human user side, which are likely to be refined after continued interaction, and after the user gets acquainted with the robot's behavior.}} \hl{Julia:leave this paragraph here or put it later in the HRI section??}

@@

The anthropomorphic form also 
increases people's expectations of the agent, and consequently, a very human-like 
appearing robot is not only perceived but also expected to act like a human. However, today's artificial systems are 
not able (yet) to well mimic human movements, or human social cues in an acceptable way. 
The case is that the artificial agent pretends to be human but fails, and this is 
more than just disappointing for the 
human -- it appears uncanny and evokes reluctance, the so-called 
\textit{"uncanny valley"}. The idea of naming this shift and drop from a positive to a negative reaction,
follows Freud's description of the so-called \textit{uncanny} (a translation from the German word 
\textit{``unheimlich''}). Controversially, something uncanny \textit{"derives its terror not from 
something externally alien or unknown but -- on the contrary -- from something 
strangely familiar which defeats our efforts to separate ourselves from it"} (cited 
after \cite{hegel_understanding_2008}).
In other words, humans are very sensitive to the limits of what is familiar to them 
in a sense of being human. At the same time, humans are naturally very skilled when 
it comes to identifying what is acceptable as a human and what is an imitation 
of it. Even subtle differences can be crucial -- humans are inherently well trained 
in identifying if there is \hl{``something strange''} in the familiar.
Summing it up, a not yet very well defined thin line separates acceptable human-like forms in 
robots from creepy artificial \textit{almost} human-like appearing robots.

@@

@@

 \begin{figure}[b]\centering
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics[scale=1]{north-wind.jpg}
  \includegraphics[scale=1]{sun.jpg}
% figure caption is below the figure
 \caption{Milo Winter illustration of Aesop Fable \textit{The North Wind and the Sun}, 1919; both natural forces wind and sun are anthropomorphized with a human face; picture source: wikipedia}
 \label{fig:north-wind-sun}       % Give a unique label
 \end{figure}

To illustrate better how anthropomorphism differs from attributions of animacy or life-
likeness, it is worth looking at paraphrases and examples of it. That is to say, 
anthropomorphism can alternatively be understood as \textbf{humanization} or 
\textbf{personification} of anything other than a human being. Personification has 
ancient roots, \eg in mythology but also in storytelling. Anthropomorphism was also 
used to describe and be able to predict natural forces or anticipate great spirits, for 
instance. An illustration of an anthropomorphism related to nature is shown in Figure 
\ref{fig:north-wind-sun}. In this Milo Winter (1919) drawing \textit{"The North Wind 
and the Sun"}, both north wind and sun are given a human face.

@@

Amongst others, one reason for this might be that there exist several understandings of what anthropomorphism is \hl{(REF Peter Ruijten, HRI 2014.)}

Literature on anthropomorphism (the social phenomenon) in robotics as well as on
anthropomorphic design of robots is quite diverse and one strives hard to
extract a coherent conclusion. One reason for this might be that robotics is a
multi-disciplinary field and researchers from very different domains might
integrate diverse or even contradictory understandings, or simply refer to
different things. However, despite the fact that there is no commonly accepted
definition, and also the terminology is not clear, the terms
\textit{anthropomorphism}, \textit{anthropomorphic} or \textit{human-like} are
often used as if their meanings were clear and agreed upon
\cite{persson_anthropomorphism_2000}. Duffy \cite{duffy_anthropomorphism_2002}
argues that these terms might even be misused (in accordance with
\cite{epley_when_2008}). For instance, some researchers refer to \textit{"the
robot's level of anthropomorphism"} \cite{bartneck_is_2007}, whereas others
argue that the system or artifact itself does not "contain anthropomorphism"
\textit{per se} but only gives rise to the process of anthropomorphizing in a
given user and situation. Consequently, Persson \textit{et al.} conclude that
anthropomorphism emerges in the \textit{interaction} between the technology and
the user \cite{persson_anthropomorphism_2000}.  \hl{[In this article, we accept
this latter viewpoint and would like to clearly distinguish between the term
\textit{anthropomorphism} as a social phenomenon and \textit{anthropomorphic} as
a specific design. This however, is not yet a definition of what exactly both
terms refer to. Still, we acknowledge that both terms can refer to different
levels or types; which we will outline later.]}

@@

@@

Anthropomorphism and similar phenomenons can also be understood as a thoughtful process of induction whereby \textit{``people reason about an unknown stimulus based on a better-known representation of a related stimulus"}~\citep{epley_when_2008}. In the case of anthropomorphism this means a person's reasoning about a non-human agent based on the representation of the self or other humans. The general assumption here is that all other agents (and entities) but oneself are \textit{``unknown''} (because we can only be our self). A central construct of this perspective is the so-called \textit{mental model} that a person has (and builds) of the agents she is reasoning about and interacting with. The specific mental model we have about an artifact / agent basically explains how this artifact / agent works the way it does. But how can we explain the phenomenon that a person uses a human-inspired mental model to explain the actions of a non-human (unknown) agent? Well, what we know and understand best is ourselves as human beings (the mental model that we have of our self and other humans). So by applying a similar mental model (or parts of it) to the unknown agent, we seek to be able to explain this unknown agent.
This view is consistent with the \emph{familiarity thesis}~\citep{hegel_understanding_2008} which claims that we understand the world based upon a mental model of the world that we are most familiar with. The authors found that human beings implicitly attribute human-like qualities, such as mental states, to nonhuman agents. This finding was evident on a neuro-physiological level as well as behavioral level. \footnote{\hl{Ha! This says that anthropomorphism can also occur on a behavioral level, yes! :)}} Also, the mental model (human-centered perspective) is supported by \cite{epley_seeing_2007} \textit{Three Factor Theory} which applies a social psychology view, and will be presented later in this section.

>>>>

Along this line, this article aims at fostering a better understanding of
\textbf{anthropomorphism} (noted \Ant in this paper) by considering the phenomenon
\textbf{as a whole}: robotics researchers often tend to believe that
anthropomorphism only describes a static set of human-like features of a robot
(like shape, speech capabilities, facial expressions, etc.). We refer to these
characteristics as the \emph{anthropomorphic design} of the
robot~\cite{fink_anthropomorphism_2012}. But \emph{anthropomorphism} in general
refers to the larger \emph{social phenomenon} that emerges from the (real or
imagined) \emph{interaction} between a robot and a
user~\citep{persson_anthropomorphism_2000}. According to~\cite{epley_when_2008},
this includes for instance emotional states, motivations, intentions
\emph{ascribed by the user} to the robot. As such, the dynamic and
socio-cognitive dimensions of anthropomorphism are essential to understand the
complex bonds that human users build with the robots.

@@

On one hand, developers and designers have discovered that human-like robots may better 
match the affordances that this to humans adapted world brings along. 
Apart from this, it is suggested that increasing the human-likeness of a robot (up to a 
certain threshold) will generally lead to 
more perceived \textit{affinity} from the human side \citep{mori_uncanny_1970} (see 
Figure \ref{fig:uncanny_valley}).
This further increases anthropomorphism, and as such
appears to be beneficial for human-robot relations, especially with socially 
interactive robots \cite{fong_survey_2003}. Consequently, the domain of (social) 
robotics tries to exploit this by using
anthropomorphic forms in robots.
Humans seem to readily accept and prefer the 
anthropomorphic form of a robot (up to a certain point), when compared to purely mechanical 
robots. A possible explanation for this is that perceiving a system as
human-like (or also pet-like) compared to machine-like changes the 
way how users try to understand and make sense of this system. The reason for this is 
that humans tend to project everyday 
\hl{expectations} of human and animalistic life onto artificial agents that use 
anthropomorphic (or zoomorphic) forms \cite{schmitz_concepts_2011}. 
Both cognitive science 
and social psychology can provide explanations for this process of projection. The basic idea 
is that \textit{"the more similar in 
appearance, the more people are likely to use themselves as a source of induction 
and anthropomorphize these nonhuman agents"} \citep{epley_seeing_2007}. Studies have further 
shown that perceived similarity between the user and the robot influences the intensity of 
the participants' attributions towards the robot. Thus,the anthropomorphic design of a robot 
impacts the user's perception of the robot, and how far the user is likely to attribute and 
experience qualities such as intelligence, fun, or intentions in the robot. However, this is not a natural spontaneous reaction to anthropomorphic form but already a thoughtful process (investigated as the process of forming a \textit{mental model}), which will be treated in more detail later, in Section \ref{sec:human-centered}).

In short, the central point in using anthropomorphic design in sociable robots is that 
humans tend to apply the same communication references to anthropomorphic robots 
that they would use with other humans, which in turn enhances feeling familiar and 
comfortable when interacting with the otherwise unknown system.
This positive effect has been shown in various 
empirical studies. It has been found that robots with human-like forms can enhance social 
responses from humans which in turn can positively impact on acceptance 
\citep{duffy_anthropomorphism_2003,goetz_cooperation_2002,venkatesh_theoretical_2000}. 
People responded more positively to an artifact that displayed human-like 
behavioral characteristics (emotions, facial expression) in contrast to a purely 
functional design 
\citep{eyssel_anthropomorphic_2010,krach_can_2008,reeves_media_1996,riek_how_2009}. 
Overall, these results suggest that human-like features in robots enhance 
anthropomorphism, and in turn create a social view toward the robot.

@@

@@

The anthropomorphic form also 
increases people's expectations of the agent, and consequently, a very human-like 
appearing robot is not only perceived but also expected to act like a human. However, today's artificial systems are 
not able (yet) to mimic well human movements, or human social cues in an acceptable way. 
The case is that the artificial agent pretends to be human but fails, and this is 
more than just disappointing for the 
human -- it appears uncanny and evokes reluctance, the so-called 
\textit{"uncanny valley"}. The idea of naming this shift and drop from a positive to a negative reaction,
follows Freud's description of the so-called \textit{uncanny} (a translation from the German word 
\textit{``unheimlich''}). Controversially, something uncanny \textit{"derives its terror not from 
something externally alien or unknown but -- on the contrary -- from something 
strangely familiar which defeats our efforts to separate ourselves from it"} (v. Foerster, cited after \cite{hegel_understanding_2008}).
In other words, humans are very sensitive to the limits of what is familiar to them 
in a sense of being human. At the same time, humans are naturally very skilled when 
it comes to identifying what is acceptable as a human and what is an imitation 
of it. Even subtle differences can be crucial -- humans are inherently well trained 
in identifying if there is \hl{``something strange''} in the familiar.
Summing it up, a not yet very well defined thin line separates acceptable human-like forms in 
robots from creepy artificial \textit{almost} human-like appearing robots.

@@

To illustrate better how anthropomorphism differs from attributions of animacy or life-
likeness, it is worth looking at paraphrases and examples of it. That is to say, 
anthropomorphism can alternatively be understood as \textbf{humanization} or 
\textbf{personification} of anything other than a human being. Personification has 
ancient roots, \eg in mythology but also in storytelling. For instance, anthropomorphism was also 
used to describe and be able to predict natural forces or anticipate great spirits.

@@

% \begin{figure}[b]\centering
%% Use the relevant command to insert your figure file.
%% For example, with the graphicx package use
%  \includegraphics[scale=1]{north-wind.jpg}
%  \includegraphics[scale=1]{sun.jpg}
%% figure caption is below the figure
% \caption{Milo Winter illustration of Aesop Fable \textit{The North Wind and the Sun}, 1919; both natural forces wind and sun are anthropomorphized with a human face; picture source: wikipedia}
% \label{fig:north-wind-sun}       % Give a unique label
% \end{figure}

%An illustration of an anthropomorphism related to nature is shown in Figure 
%\ref{fig:north-wind-sun}. In this Milo Winter (1919) drawing \textit{"The North Wind 
%and the Sun"}, both north wind and sun are given a human face.

@@

Literature on anthropomorphism (the social phenomenon) in robotics as well as on
anthropomorphic design of robots is quite diverse. There exist different understandings of what anthropomorphism is, and consequently, one strives hard to
extract a coherent conclusion.

One reason for why anthropomorphism is defined in different ways might be that it
is a concept that is studied in very different domains that might each
integrate diverse or even contradictory understandings, or simply refer to
different things.
However, one conflict is that, despite the fact that there is no commonly accepted
definition, and also the terminology is not clear, the terms
\textit{anthropomorphism}, \textit{anthropomorphic} or \textit{human-like} are
often used as if their meanings were clear and agreed upon
\cite{persson_anthropomorphism_2000}. Duffy \cite{duffy_anthropomorphism_2002}
argues that these terms might even be misused (in accordance with
\cite{epley_when_2008}). For instance, some researchers refer to \textit{"the
robot's level of anthropomorphism"} \cite{bartneck_is_2007}, whereas others
argue that the system or artifact itself does not "contain anthropomorphism"
\textit{per se} but only gives rise to the process of anthropomorphizing in a
given user and situation. Consequently, Persson \textit{et al.} conclude that
anthropomorphism emerges in the \textit{interaction} between the technology and
the user \cite{persson_anthropomorphism_2000}. We will come back to this view later, in the summary of this section (Subsection \ref{sec:multi-factors}).

@@

Similarly \cite{ruijten_introducing_2014} argues that anthropomorphism

@@

Anthropomorphism and similar phenomenons can also be understood as a thoughtful process of induction whereby \textit{``people reason about an unknown stimulus based on a better-known representation of a related stimulus"}~\citep{epley_when_2008}. In the case of anthropomorphism this means a person's reasoning about a non-human agent based on the representation of the self or other humans. A central construct of this perspective is the so-called \textit{mental model} that a person has (and builds) of the agents she is reasoning about and interacting with. The specific mental model we have about an artifact / agent basically explains (in our individual way to ourselves) how this artifact / agent works the way it does.
There are two central aspects here, when looking at the quote from \citep{epley_when_2008} above.
\textit{(1)} First, one assumption is that the user is not familiar with the other agent. This is basically true for all other agents (and entities) but oneself because we can only be our self. This view is also consistent with the \emph{familiarity thesis}~\citep{hegel_understanding_2008} which claims that we understand the world based upon a mental model of the world that we are most familiar with (our self). Thus, we can best use our self as source of induction when reasoning about other agents. However, in case we would like to understand the behavior of a spider, we would probably not use our mental model of a human (or our self) simply because we do not have many things in common with a spider. And this is the second central aspect here.
\textit{(2)} Second, one assumption is based oj the anthropomorphic design of the other agent (see Section \ref{sec:anthropomorphic-design}): if the agent behaves (appears) much like a human being (\eg a robot that emits a human voice), people's mental model of the agent's behavior is likely to approach their mental model of humans (though the model may differ in some important aspects).

This cognitive viewpoint on anthropomorphism is important because people's estimation of an agent's knowledge model and its capabilities affects the way they relate to it. This holds important implications for the resulting interaction with the system, the user experience, and the acceptance. Previous research examined the validity of the mental model concept with various kinds of robots \cite{schmitz_concepts_2011,kiesler_mental_2002}. Findings suggest that people tend to hold richer mental models of anthropomorphic robots in contrast to mechanic ones \cite{kiesler_mental_2002}.\footnote{\hl{Later in this article, we may however come to question how far this result is based on expectations from the human user side, which are likely to be refined after continued interaction, and after the user gets acquainted with the robot's behavior.}} A similar finding is described in \cite{hegel_understanding_2008}. In a user study, the authors found that participants implicitly attributed human-like qualities, such as mental states, to their non-human interaction partner. This finding was evident on a neuro-physiological level as well as behavioral level. \footnote{\hl{Ha! This says that anthropomorphism can also occur on a behavioral level, yes! :)}}
